Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-09-28 prop_letter.tex
prop_letter.tex \section{Gleason} Merges Prof. Hanson's suggestion after meeting...

While the finite-precision extension of the Kochen-Specker theorem
asserts the non-existence of a certain kind of QIVPMs, we also want
to understand what a QIVPM could be when it exists. A conventional
quantum probability measure~$\mu$ can easily be constructed using
the Born rule when one knows the current state of the system~\citep{Born1983,peres1995quantum,544199,Jaeger2007}.
Gleason's theorem asserts that the only state consistent with $\mu$
on all projectors is the one used to construct $\mu$~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.
In other words, the set of all quantum probability measures corresponds
one-to-one to the set of states so that studying the set of states
is sufficient to understand all possible phenomenon regarding to quantum
probability. Although we can extend the Born rule to its finite precision
version easily, there are some subtleties extending Gleason's theorem.
If we only focus on the Born QIVPMs~$\bar{\mu}$, that is, QIVPM's
induced by a state~$\rho$ according to the Born rule, we find that
if we monotonically shrink the length of intervals, the states consistent
with $\bar{\mu}$ will be closer and closer to $\rho$, and vice versa.
However, there is a QIVPM such that no state can be consistent with
it on every projector. In another word, although the Born rule and
Gleason's theorem can be extended to consider finite precision measurement,
the problem of finite precision measurement itself is not limited
in just applying error analysis in experimental results, but it highlights
some new phenomenon regarding to the foundation of quantum machines,
and it is necessary to embedding the idea of the resources available
for realistic measurement into the formulation of quantum theory.

The first step toward a finite precision Born rule is to find a way
of replacing the computed probability by an interval. This can be
done by carefully picking a set of intervals and a rule to select
the correct interval when intervals are overlapping. For example,
consider the following collection of overlapping intervals: 
\begin{equation}
\begin{aligned}\mathscr{I}_{n}= & \set{\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\\
 & \cup\{\imposs,\necess\}
\end{aligned}
\end{equation}
which are overlapping intervals with length~$1/n$ except for the
intervals with length~$\left(2n\right)^{-1}$ at each end. As in
Table~\ref{tab:I_n}, to select the correct interval from these overlapping
intervals~$[\ell,r]$, we define the non-overlapping interior part
of a interval, $\textrm{Int}\left([\ell,r]\right)$.
\begin{table}
\caption{\label{tab:I_n}Intervals in $\mathscr{I}_{n}$, their lengths, and
their interiors, where $2\mid k$ means $k$ is even, and $2\nmid k$
means $k$ is odd. Although intervals~$[\ell,r]$ are overlapping,
their interior~$\textrm{Int}\left([\ell,r]\right)$ are non-overlapping.
See Table~\ref{tab:I_2} for a concrete example when $n=2$.}

@@

\noindent Since $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$ clearly satisfies
Eqs.~(\ref{eq:QIVPM-constraints}), the condition really needing
to be checked is Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}).
Given $x_{0}$, $x_{1}$, and $x_{0}+x_{1}\in\left[0,1\right]$, it
is easy to verify that
\begin{equation}
\iota_{n}\left(x_{0}+x_{1}\right)\subseteq\iota_{n}\left(x_{0}\right)+\iota_{n}\left(x_{1}\right)\,.\label{eq:iota-add}
\end{equation}
This implies Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
when $P_{0}$ and $P_{1}$ are orthogonal. When $P_{0}$ and $P_{1}$
are not orthogonal on a three-dimensional Hilbert space, we have two
possibilities:
\begin{itemize}
\item Their complements are orthogonal so that we can apply Eq.~(\ref{eq:iota-add})
on their complements.
\item One of them projects onto a subspace of what the other one projects
onto.
\end{itemize}
\noindent No matter which case Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
can be deduced easily. Therefore, $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
is indeed a QIVPM.

@@

\begin{thm}[$\mathscr{I}_{\infty}$ Variant of Gleason]\label{cor:Gleason's}In
a Hilbert space $\Hilb$ of dimension $d\geq3$, given a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\infty}$,
the states consistent with it on every projector, $\coreBorn\left(\bar{\mu},\events\right)$,
has a unique element. \end{thm}

>>>>

While the finite-precision extension of the Kochen-Specker theorem
asserts the non-existence of a certain kind of QIVPMs, we also want
to understand what a QIVPM could be when it exists. A conventional
quantum probability measure~$\mu$ can easily be constructed using
the Born rule when one knows the current state of the system~\citep{Born1983,peres1995quantum,544199,Jaeger2007}.
Gleason's theorem asserts that the only state consistent with $\mu$
on all projectors is the one used to construct $\mu$ by the Born
rule~\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}. In
other words, the set of all quantum probability measures corresponds
one-to-one to the set of states, so that studying the set of states
is sufficient to understand all possible phenomena regarding to quantum
probability. Although we can extend the Born rule to its finite precision
version easily, there are some subtleties to extending Gleason's theorem.
If we only focus on the Born QIVPMs~$\bar{\mu}$, that is, QIVPM's
induced by a state~$\rho$ according to the Born rule, we find that
if we monotonically shrink the length of intervals, the states consistent
with $\bar{\mu}$ will be closer and closer to $\rho$, and vice versa.
However, there is a QIVPM such that no state can be consistent with
it on every projector. In another words, although the Born rule and
Gleason's theorem can be extended to consider finite precision measurement,
the problem of finite precision measurement itself is not limited
to error analysis in experimental results, but it highlights some
new phenomena regarding the foundation of quantum mechanics, and it
is necessary to embed the idea of the resources available for realistic
measurement into the formulation of quantum theory.

The first step toward a finite precision Born rule is to find a way
of replacing the computed probability by an interval. This can be
done by carefully picking a set of intervals and a rule to select
the correct interval when intervals are overlapping. For example,
consider the following collection of overlapping intervals: 
\begin{equation}
\begin{aligned}\mathscr{I}_{n}= & \set{\left[\tfrac{k-1}{2n},\tfrac{k+1}{2n}\right]\cap\left[0,1\right]}{k=0,\ldots,2n}\\
 & \cup\{\imposs,\necess\}
\end{aligned}
\end{equation}
which are overlapping intervals with length~$1/n$ except for the
intervals with length~$\left(2n\right)^{-1}$ at each end. As shown
in Table~\ref{tab:I_n}, we select the correct interval from these
overlapping intervals~$[\ell,r]$, and define $\textrm{Int}\left([\ell,r]\right)$,
the non-overlapping interior part of an interval.
\begin{table}
\caption{\label{tab:I_n}Intervals in $\mathscr{I}_{n}$, their lengths, and
their interiors, where $2\mid k$ means $k$ is even, and $2\nmid k$
means $k$ is odd. Although intervals~$[\ell,r]$ are overlapping,
their interior~$\textrm{Int}\left([\ell,r]\right)$ are non-overlapping.
See Table~\ref{tab:I_2} for a concrete example when $n=2$.}

@@

\noindent Since $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$ clearly satisfies
Eqs.~(\ref{eq:QIVPM-constraints}), the condition that needs to be
checked is Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion}).
Given $x_{0}$, $x_{1}$, and $x_{0}+x_{1}\in\left[0,1\right]$, it
is easy to verify that
\begin{equation}
\iota_{n}\left(x_{0}+x_{1}\right)\subseteq\iota_{n}\left(x_{0}\right)+\iota_{n}\left(x_{1}\right)\,.\label{eq:iota-add}
\end{equation}
This implies Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
when $P_{0}$ and $P_{1}$ are orthogonal. When $P_{0}$ and $P_{1}$
are not orthogonal on a three-dimensional Hilbert space, we have two
possibilities:
\begin{itemize}
\item Their complements are orthogonal so that we can apply Eq.~(\ref{eq:iota-add})
on their complements.
\item One of them projects onto a subspace of what the other one projects
onto.
\end{itemize}
\noindent No matter which case holds, Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
can be deduced easily. Therefore, $\iota_{n}\circ\mu_{\rho}^{\mathrm{B}}$
is indeed a QIVPM.

@@

\begin{thm}[$\mathscr{I}_{\infty}$ Variant of Gleason]\label{cor:Gleason's}In
a Hilbert space $\Hilb$ of dimension $d\geq3$, given a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\infty}$,
the state consistent with it on every projector, $\coreBorn\left(\bar{\mu},\events\right)$,
is unique. \end{thm}
